Quantizing the Bosonic String on a Loop Quantum Gravity Background

### 2024-07-09 lqg-strings.tex
Added to discussion section

In this article we have presented a phenomenological modification of the Nambu-Goto action, inspired on the one hand, by the central result of loop quantum gravity, according to which area and other geometric observables are quantized at the Planck scale. In particular the spectrum of the area operator is discrete and the smallest eigenvalue is non-zero (of the order $l_{pl}^2 \sim 10^{-68}\,\text{m}^2$). On the other hand we also took as a working hypothesis the statement that string theory, or some suitable version of it, provides the correct description of the low energy physics of the world around us. We then searched for the way in which both of these paradigms can be incorporated into a single theoretical model. The form of the action shown in \eqref{eqn:ng-area-corrected} appears, at least superficially, as the simplest possible modification of the Nambu-Goto action which incorporates a minimum eigenvalue for the area. This value is encoded in the modulus of the $\Delta(x)$ field. In the process we introduced a new coupling parameter $ k = l_{pl}/l_s$ which is the ratio of the string scale to the Planck scale.

@@

Let us make one crucial observation here. In the expression \eqref{eqn:ng-area-corrected} we originally took $\Delta$ to a constant representing the smallest quantum of area. It is only \emph{later}, by demanding consistency of the area-corrected Nambu-Goto and Polaykov actions that we realise that $\Delta \equiv \Delta_{ab}(x) $ is in fact an antisymmetric worldsheet two-form.

Let us recall the steps which have led us here. Our original motivation was to continue to expand on the work which was started by one of us several years ago \cite{Vaid2017Connecting} and ask whether it is possible to construct an alternative to the usual action for the bosonic string in which the 2d geometry of the string worldsheet is coupled to the bulk geometry, not through the bulk metric $G_{\mu\nu}$, but through the pullback of a bulk connection to the worldsheet. This allows us to define a covariant derivative on the worldsheet which acts on the embedding co-ordinate fields $X^\mu$

@@

\begin{itemize}
	\item We adopt the central result of LQG which is quantization of area. 
	\item We seek to describe the behavior of a string in an arbitrarily curved background geometry.
	\item Connection variables are used in LQG to describe geometry. So we should search for non-metrical description of coupling string geometry to bulk.
	\item For arbitrarily curved bulk geometry, global Poincare symmetry of embedding fields is meaningless.
	\item Elevate global Poincare to local Poincare following the lessons of gauge theory, which leads to introduction of extra "holographic" co-ordinate.
	\item Investigate physical implications of holographic co-ordinate for string phenomenology in this picture.
\end{itemize}

@@

We want to conclude this discussion by pointing out some complementary aspects of the historical development of both string theory and loop quantum gravity which should already have been a hint many years ago that the two would eventually turn out to be intimately related.

What this very brief history of the early years of string theory tells us is that the theory arose from trying to understand particle scattering. This, in part, explains why string theorists didn't pay much attention to the question of background independence of string theory during its early decades. For particle physicists the flat Minkowski background was a given and so it was natural for them to begin to quantize the string by assuming the flat Minkowski metric for the bulk geometry.

>>>>

In this article we have presented a phenomenological modification of the Nambu-Goto action \emph{inspired by the central result of loop quantum gravity, according to which area and other geometric observables are quantized at the Planck scale}. In particular the spectrum of the area operator is discrete and the smallest eigenvalue is non-zero (of the order $l_{pl}^2 \sim 10^{-68}\,\text{m}^2$). On the other hand we also took as a working hypothesis the statement that string theory, or some suitable version of it, provides the correct description of the low energy physics of the world around us. We then searched for the way in which both of these paradigms can be incorporated into a single theoretical model. The form of the action shown in \eqref{eqn:ng-area-corrected} appears, at least superficially, as the simplest possible modification of the Nambu-Goto action which incorporates a minimum eigenvalue for the area. This value is encoded in the modulus of the $\Delta(x)$ field. In the process we introduced a new coupling parameter $ k = l_{pl}/l_s$ which is the ratio of the string scale to the Planck scale.

@@

Let us make one crucial observation here. In the expression \eqref{eqn:ng-area-corrected} we originally took $\Delta$ to a constant representing the smallest quantum of area. It is only \emph{later}, by demanding consistency of the area-corrected Nambu-Goto and Polaykov actions that we realise that $\Delta \equiv \Delta_{ab}(x) $ is in fact an antisymmetric worldsheet two-form!

We would also like to inform the discussion by pointing out some complementary aspects of the historical development of both string theory and loop quantum gravity which should already have been a hint many years ago that the two would eventually turn out to be intimately related.

What this very brief history of the early years of string theory tells us is that the theory arose from trying to understand particle scattering. This, in part, explains why string theorists didn't pay much attention to the question of background independence of string theory during its early decades. For particle physicists the flat Minkowski background was a given - the entire machinery of Feynman diagrams and perturbation theory was built with the implicit assumption of a flat background manifold. The ``string'' in the first conception was solely meant to be a description of bound states of mesons and hadrons. From the abstract of Susskind's 1970 paper \cite{Susskind1970Dual} we quote

\begin{quote}
	\emph{A theory of mesons is based on the experimentally observed linear level spacing of hadron excitations. The theory uses a model for the internal structure of mesons consisting of a system of harmonic vibrations. Roughly speaking, a meson is described by the degrees of freedom of a four-dimensional rubber band with a quark pair embedded in it.}
\end{quote}

It was realized only later \cite{Scherk1974bDual} that excitations of the string included a spin-2 object which could be identified with a graviton.

@@

\begin{itemize}
	\item We seek to describe the behavior of a string in an arbitrarily curved background geometry.
	\item Connection variables are used in LQG to describe geometry. So we should search for non-metrical description of coupling string geometry to bulk.
	\item For arbitrarily curved bulk geometry, global Poincare symmetry of embedding fields is meaningless.
	\item Elevate global Poincare to local Poincare following the lessons of gauge theory, which leads to introduction of extra "holographic" co-ordinate.
	\item Investigate physical implications of holographic co-ordinate for string phenomenology in this picture.
\end{itemize}

@@
